noam_data_exploration_v1

### 2024-07-17 noam_data_exploration_v3.Rmd
Figured out how to make portfolio effects work at quadrat level, but the data are weird because of too many quadrats with extremely low richness

problem_children <- c("LTER 5 Fringing Reef Algae Transect 3 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 8", "LTER 5 Fringing Reef Algae Transect 3 Quad 6", "LTER 5 Fringing Reef Algae Transect 3 Quad 3", "LTER 5 Fringing Reef Algae Transect 3 Quad 10")

@@

for(i in 1:length(unique(unique(data$Location))) ){ # for each  quadrat
  # create a new dataframe for current quadrat
  newData <- subset(data, Location == levels(data$Location)[i])
  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
  if(sum(current_matrix) == 0){ # if this was a 0 algae quadrat
    quadrat_level_portfolio[i, 2] <- NA # set portfolio effects to 0
  }else{ # otherwise
    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae

@@

>>>>

problem_children <- c("LTER 5 Fringing Reef Algae Transect 2 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 9", "LTER 5 Fringing Reef Algae Transect 3 Quad 8", "LTER 5 Fringing Reef Algae Transect 3 Quad 6", "LTER 5 Fringing Reef Algae Transect 3 Quad 3", "LTER 5 Fringing Reef Algae Transect 3 Quad 10")

@@

for(i in 1:length(unique(unique(data$Location))) ){ # for each  quadrat
  # create a new dataframe for current quadrat
  newData <- subset(data, Location == levels(data$Location)[i])
  current_matrix <- newData[ , -which(names(newData) %in% meta_cols)] # remove metadata columns
  if(sum(current_matrix) == 0 | # if this was a 0 algae quadrat
     # or if there was only 1 column after removing all irrelevant algae
     is.null(ncol(current_matrix[, which(colSums(current_matrix) != 0)]))){ 
    quadrat_level_portfolio[i, 2] <- NA # set portfolio effects to 0
  }else{ # otherwise
    current_matrix <- current_matrix[, which(colSums(current_matrix) != 0)] # remove irrelevant algae

@@

sum(is.na(quadrat_level_portfolio$Portfolio_effect)) # 4 quadrats causing trouble
problem_children <- droplevels(quadrat_level_portfolio[is.na(quadrat_level_portfolio$Portfolio_effect),]$Location)
hist(quadrat_level_portfolio$Portfolio_effect) # some crazy outliers due to low richness
